---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

1. Load the data (i.e. `read.csv()`): 

```{r echo = TRUE}
activity <- read.csv(file = "activity.csv")
```

Show the first lines: 
```{r echo = TRUE}
head(activity)
```

Show the class of variables: 
```{r echo = TRUE}
str(activity)
```

Convert date from char to date class: 
```{r echo = TRUE}
activity$date <- as.Date(activity$date, "%Y-%m-%d")
str(activity)

```

## What is mean total number of steps taken per day?

1. Make a histogram of the total number of steps taken each day

Use dplyr functions to group data by date and sum the steps by day:
```{r echo = TRUE}
library(dplyr)
steps_by_date <- activity %>%
        group_by(date) %>%
        summarise(steps = sum(steps))

head(steps_by_date)
```

Plot an histogram of number of steps by day:

```{r echo = TRUE}
library(ggplot2)
g <- ggplot(data = steps_by_date, aes(x = date, y = steps))
g + geom_bar(stat = 'identity')

```

2. Calculate and report the **mean** and **median** total number of steps taken per day

Mean: 
```{r echo = TRUE}
mean_steps_by_date <- activity %>%
        group_by(date) %>%
        summarise(mean_steps = mean(steps, na.rm = TRUE))

steps_by_date <- mutate(steps_by_date, mean_steps = mean_steps_by_date$mean_steps)

head(steps_by_date)
```

Median: 
```{r echo = TRUE}
median_steps_by_date <- activity %>%
        group_by(date) %>%
        summarise(median_steps = median(steps, na.rm = TRUE))

steps_by_date <- mutate(steps_by_date, median_steps = median_steps_by_date$median_steps)

head(steps_by_date)
```
```{r}
mean(steps_by_date$steps, na.rm = TRUE)
median(steps_by_date$steps, na.rm = TRUE)
```


## What is the average daily activity pattern?

1. Make a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r echo = TRUE}
steps_by_interval <- activity %>%
        group_by(interval) %>%
        summarise(mean_steps = mean(steps, na.rm = TRUE))

qplot(x = interval, y = mean_steps, data = steps_by_interval, geom = "line", xlab = "5-min intervals", ylab = "Steps averaged across all days")
```
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r echo = TRUE}
temp <- activity %>%
        group_by(interval) %>%
        summarise(max_steps = max(steps, na.rm = TRUE))

steps_by_interval <- steps_by_interval %>%
        mutate(temp)

qplot(x = interval, y = max_steps, data = temp, geom = "line", xlab = "5-min intervals", ylab = "Maximum steps across all days")
```

## Imputing missing values  

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with `NA`s)


```{r echo = TRUE}

sapply(activity, function(x) sum(is.na(x)))

```
2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

Duplicate steps variable to "imputed_steps" variable

```{r echo = TRUE}
activity <- activity %>%
        mutate (steps_imputed = steps)

summary(activity)

```
Replace NA values by mean values of the 5-min interval :
```{r echo = TRUE}

activity <- merge(activity, steps_by_interval, by.x = "interval", by.y = "interval", all = FALSE)

activity$steps_imputed[is.na(activity$steps)] <- activity$mean_steps[is.na(activity$steps)]

```
3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r echo = TRUE}

activity_imputed <- activity %>%
        select("interval", "steps_imputed", "date")

names(activity_imputed) <- c("interval", "steps", "date")

summary(activity_imputed)

```

```{r echo = TRUE}

summary(activity)
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the **mean** and **median** total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?


Histogram of steps per date with imputation:
```{r echo = TRUE}

steps_by_date_imputed <- activity_imputed %>%
        group_by(date) %>%
        summarise(steps = sum(steps))

g <- ggplot(data = steps_by_date_imputed, aes(x = date, y = steps))
g + geom_bar(stat = 'identity')
```

```{r echo=TRUE}

mean(steps_by_date_imputed$steps)
median(steps_by_date_imputed$steps)
```
The **mean** value is the **same** as the value before imputing missing data because we put the mean value for that particular 5-min interval. The median value shows **a little** difference.

## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}

weekdays_names <- c('lundi', 'mardi', 'mercredi', 'jeudi', 'vendredi')

activity_imputed <- activity_imputed %>%
  mutate(day = weekdays(date)) %>%
  mutate(factor = factor((weekdays(date) %in% weekdays_names), 
         levels=c(FALSE, TRUE), labels=c('weekend', 'weekday'))) 


```


2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged
across all weekday days or weekend days (y-axis). The plot should look something like the following, which was creating using simulated data:

```{r}
steps_by_interval_imputed <- activity_imputed %>%
        group_by(interval, factor) %>%
        summarise(steps = mean(steps))



ggplot(steps_by_date_imputed, aes(x = date, y = steps)) + geom_line() + facet_grid(day)


```







